\documentclass[a4paper,12pt]{book}
\usepackage[utf8]{inputenc}
\title{}
\author{Rachel Morris}
\date{\today}

\usepackage{rachwidgets}
\usepackage{fancyhdr}
\usepackage{lastpage}
\usepackage{dirtree}
\usepackage{boxedminipage}

\setcounter{chapter}{6}
\setcounter{section}{4}
\newcommand{\laChapter}{6.5 Recursion Revisited\ }

\newcommand{\laClass}{CS 211\ }
\newcommand{\laSemester}{Fall 2017\ }
\newcounter{question}

\pagestyle{fancy}
\fancyhf{}
\lhead{CS 211 Exercise}
\chead{Fall 2017}
\rhead{Ch \laChapter}
\rfoot{\thepage\ of \pageref{LastPage}}
\lfoot{\scriptsize Compiled by Rachel Morris, last updated \today}

\renewcommand{\headrulewidth}{2pt}
\renewcommand{\footrulewidth}{1pt}

\begin{document}

    %\toggletrue{answerkey}
    \togglefalse{answerkey}
    
    \section*{Running trials in code}

        Let's write a recursive function to run these types of trials
        for us and get a simulated average value.

        ~\\
        In Visual Studio create a new \textbf{Empty Project}, and add
        one source file to it. Make sure to include the following libraries:

\begin{verbatim}
#include <iostream>     // output
#include <cstdlib>      // random
#include <ctime>        // time
using namespace std;    // standard library
\end{verbatim}

        Start with the following shell function:
        
\begin{verbatim}
int RunTrial( int min, int max, int stopAtFirst, int trialCount )
{
}
\end{verbatim}

        And main():

\begin{verbatim}
int main()
{
    srand( time( NULL ) );
    int min, max, stopAtFirst, count;
    
    return 0;
}
\end{verbatim}

\newpage

\textbf{Starter code}

\begin{lstlisting}[style=code]
#include <iostream>     // output
#include <cstdlib>      // random
#include <ctime>        // time
using namespace std;    // standard library

int RunTrial(   int min, int max,
                int stopAtFirst, int trialCount )
{
}

int main()
{
    // seed the random # generator
    srand( time( NULL ) );  
    int min, max, stopAtFirst, count;
    
    return 0;
}
\end{lstlisting}
    ~\\
    
    In C++, you can display text to the screen with a \texttt{cout} statement...

\begin{verbatim}
cout << "Message";
\end{verbatim}

    And you can get input with a \texttt{cin} statement, storing it in a variable...

\begin{verbatim}
cin >> min;
\end{verbatim}

    \textbf{Get user input:}
    Ask the user to enter the \textbf{Minimum value}, \textbf{Maximum value},
    \textbf{Which value to stop at}, and \textbf{How many times to run it}.
    Store these in the variables declared at the top of main().

    ~\\
    When running the program, the user can enter the following to
    simulate a die roll or a coin flip:

    \begin{center}
        \begin{tabular}{ l c c }
            & min & max
            \\ \hline
            Die & 1 & 6
            \\
            Coin & 1 & 2
        \end{tabular}
    \end{center}

    \newpage


    \textbf{Total trials:}
    Create an integer variable that will store
    the total amount of trials that were run across all experiments.
    Make sure to initialize it to 0.
    ~\\
    
    \textbf{Experiment run:}
    After getting the user's input, you will need to write a for-loop
    to run the experiment the specified amount of times.
    ~\\

    \textbf{Calling the function:}
    Within the for-loop, you will call the \texttt{RunTrial} function.
    The output of this function is the total amount of trials that
    were ran before getting the specified value for the first time.
    You will add this value onto the \textbf{totalTrials} variable.
    ~\\
    
    \textbf{Calculating the average:}
    Once the experiments has completed (the for-loop is over), you will
    calculate the average amount of trials that were ran.
    Create a float variable called \textbf{average} and divide
    the \textbf{totalTrials} by the \textbf{count}.
    Make sure you're doing float division or it won't turn out correctly.

    ~\\

    See the next page for the main() code so far.


\newpage
\begin{lstlisting}[style=code]
int main()
{
    // seed the random # generator
    srand( time( NULL ) );  
    int min, max, stopAtFirst, count;
    
    cout << "Minimum value: ";
    cin >> min;

    cout << "Maximum value: ";
    cin >> max;

    cout << "Stop at what: ";
    cin >> stopAtFirst;
    
    cout << "How many times to run: ";
    cin >> count;

    int totalTrials = 0;

    for ( int i = 0; i < count; i++ )
    {
        cout << endl;
        cout << "Running trial set " << i << "... \t";
        int trials = RunTrial( min, max,
            stopAtFirst, 1 );
        totalTrials += trials;
        cout << "\t" << trials << " trials ran";
    }
    
    float averageTrials =
        float( totalTrials ) / float( count );
        
    cout << "Average trials ran: " << averageTrials
        << endl << endl;
    
    return 0;
}
\end{lstlisting}

    \newpage

    Now to work on the recursive function.
    
\begin{verbatim}
int RunTrial( int min, int max, int stopAtFirst, int trialCount )
{
}
\end{verbatim}

    A \textbf{Recursive Function} is a function that will call itself
    until the job is done. It is useful for breaking a problem into pieces.

    At the top of this function, you will get a random value with the following:

\begin{verbatim}
// Get a random value between MIN and MAX, inclusive
int diff = (max + 1 - min);
int n = rand() % diff + min;
cout << n << " ";
\end{verbatim}

    Recursive functions need a \textbf{Base-case}, which is the scenario
    where it will end. In our case, the function stops once we get
    the \textbf{stopAtFirst} value.
    Use an if statement to check if the value of \textbf{n} is equal
    to \textbf{stopAtFirst}. If it is, return the current value of the
    \textbf{trialCount}.

    Otherwise, we will call the recursive function.
    All the variables are the same, except we add 1 to the trial count.
    
\begin{verbatim}
return RunTrial( min, max, stopAtFirst, trialCount + 1 );
\end{verbatim}

    Altogether, it will look like this:

\begin{lstlisting}[style=code]
int RunTrial( int min, int max,
    int stopAtFirst, int trialCount )
{
    int diff = (max + 1 - min);
    int n = rand() % diff + min;

    cout << n << " ";
    
    if ( n == stopAtFirst ) // Base case
    {
        return trialCount;
    }

    return RunTrial( min, max,
        stopAtFirst, trialCount + 1 );
}
\end{lstlisting}

\newpage

    When the program is run, we can execute the experiment a large
    amount of times to see if the average that comes out is close
    to the expected value we calculate.
    
\begin{lstlisting}[style=output]
Minimum value: 1
Maximum value: 2
Stop at what: 1
How many times to run: 10000

(...)

Running trial set 9989... 	1 	1 trials ran
Running trial set 9990... 	1 	1 trials ran
Running trial set 9991... 	1 	1 trials ran
Running trial set 9992... 	1 	1 trials ran
Running trial set 9993... 	1 	1 trials ran
Running trial set 9994... 	2 2 2 1 	4 trials ran
Running trial set 9995... 	2 1 	2 trials ran
Running trial set 9996... 	2 2 1 	3 trials ran
Running trial set 9997... 	2 1 	2 trials ran
Running trial set 9998... 	2 2 1 	3 trials ran
Running trial set 9999... 	2 1 	2 trials ranAverage trials ran: 2.0233
\end{lstlisting}

From the example above, our $E[X]$ was 2, so this is pretty close.

\end{document}
