\input{BASE-HEAD}
\newcommand{\laClass}       {CS 211}
\newcommand{\laSemester}    {Spring 2018}
\newcommand{\laChapter}     {6.2}
\newcommand{\laType}        {Exercise}
\newcommand{\laPoints}      {6}
\newcommand{\laTitle}       {Sum and Product Rules for Probability}
\newcommand{\laDate}        {Feb 22, 2018}
\setcounter{chapter}{5}
\setcounter{section}{1}
\addtocounter{section}{-1}
\newcounter{question}

\toggletrue{answerkey}
\togglefalse{answerkey}

\input{BASE-HEADER}

\input{INSTRUCTIONS-EXERCISE}

\notonkey{
% ASSIGNMENT ------------------------------------ %

    \section{\laTitle}

    \subsection{The Sum Rule}
        \begin{intro}{Disjoint events}
            Two events are said to be \textbf{disjoint} (or \textit{mutually exclusive}
            if they cannot occur simultaneously.
            \footnote{From Discrete Math by Ensley and Crawley, page 448}
        \end{intro}

    \stepcounter{question}
    \begin{questionNOGRADE}{\thequestion}
        For each of the experiments given below, decide if the
            events described are disjoint:

            \begin{itemize}
                \item[a.]	When tossing a coin four times, let $E_1$ be the
                event that there are exactly three heads and $E_2$ be the event
                that there are exactly two heads. ~\\ \vspace{1cm}

                \item[b.]	When choosing four cards, let $E_1$ be the event that
                the cards have the same value and $E_2$ be the event that the cards
                have the same suit. ~\\  \vspace{1cm}

                \item[c.] When choosing a committee of three people from a
                club with 8 men and 12 women, let $E_1$ be the event that the committee
                has a woman and let $E_2$ be the event that the committee has a man. ~\\
            \end{itemize}        
    \end{questionNOGRADE}

    \newpage

        \begin{intro}{The Sum Rule, Theorem 1}
            If $E_1$ and $E_2$ are disjoint events in a given experiment,
            then the probability that $E_1$ or $E_2$ occurs is the sum of
            $Prob(E_1)$ and $Prob(E_2)$. That is,

            $$Prob(E_1 or E_2) = Prob(E_1) + Prob(E_2)$$

            for disjoint events.
            \footnote{From Discrete Math by Ensley and Crawley, page 449}
        \end{intro}

    \stepcounter{question}
    \begin{questionNOGRADE}{\thequestion}
        A card is drawn from an ordinary deck of 52 cards. Show how
        to use the basic sum rule to find the probability that
        the card is...
        
        \begin{itemize}
            \item[a.] An ace or a jack.
                    \begin{hint}{Hint}
                        $E_{1}$ is the set of outcomes where you get an Ace... \\
                        \{ Ace-Heart, Ace-Diamond, Ace-Spade, Ace-Club \}
                        and the probability of getting an Ace, $Prob(E_{1})$
                        is 4 outcomes out of 52, or $\frac{1}{13}$.
                    \end{hint}
                    ~\\ \vspace{1cm}

            \item[b.] A diamond or a black jack, queen, or king card.
                    ~\\ \vspace{2cm}

            \item[c.] An even number value or a red jack, queen, or king card.
                    ~\\ \vspace{1cm}
        \end{itemize}
    \end{questionNOGRADE}

    \newpage

        \begin{intro}{The General Sum Rule, Theorem 2}
            If $E_1$ and $E_2$ are any events in a given experiment,
            then the probability that $E_1$ or $E_2$ occurs is given by
            $$ Prob(E_1 or E_2) = Prob(E_1) + Prob(E_2) - Prob(E_1 and E_2) $$

            If $E_1$ and $E_2$ are disjoint, then $E_1 \cap E_2 = \emptyset$,
            so $Prob(E_1 and E_2) = 0$.
            \footnote{From Discrete Math by Ensley and Crawley, page 450}
        \end{intro}

    \stepcounter{question}
    \begin{questionNOGRADE}{\thequestion}
        % Exercise 7
        What is the probability that when a pair of dice are rolled,
        either (at least) one die shows a 5 or the dice sum to 8?

            \begin{center}
                \footnotesize
                \begin{tabular}{c c c c c c}
                    (1,1) & (1,2) & (1,3) & (1,4) & (1,5) & (1,6) \\
                        (2,1) & (2,2) & (2,3) & (2,4) & (2,5) & (2,6) \\
                    (3,1) & (3,2) & (3,3) & (3,4) & (3,5) & (3,6) \\
                        (4,1) & (4,2) & (4,3) & (4,4) & (4,5) & (4,6) \\
                    (5,1) & (5,2) & (5,3) & (5,4) & (5,5) & (5,6) \\
                        (6,1) & (6,2) & (6,3) & (6,4) & (6,5) & (6,6)
                \end{tabular}
            \end{center}

        \begin{itemize}
            \item[a.]   What is the sample size? $n(S) = $

            \item[b.]   How many outcomes $n(E_1)$ are there where you get either 5 for the first die, 5 for the second die, or 5 for both dice? Write out the set of $E_1$.
                        ~\\ \vspace{.5cm}

            \item[c.]   How many outcomes $n(E_2)$ are there where the two dice sum to 8? Write out the set of $E_2$.
                        ~\\ \vspace{.5cm}

            \item[d.]   What is the amount of overlap $n(E_1 \tab[0.5cm]AND\tab[0.5cm] E_2)$? Write out this set.
                        ~\\ \vspace{.5cm}

            \item[e.]   Use the General Sum Rule to find the probability that you will get either at least one die showing a 5, OR the dice sum to 8.
                        ~\\
        \end{itemize}

    \end{questionNOGRADE}

    \subsection{The Product Rule}
        \begin{intro}{Independent events}
            Two events are said to be \textbf{independent} if the occurrence
            of one event is not influenced by the occurred (or nonoccurrence)
            of the other event.
            \footnote{From Discrete Math by Ensley and Crawley, page 451}
        \end{intro}
    
    \stepcounter{question}
    \begin{questionNOGRADE}{\thequestion}
        For each of the following experiments given below,
        decide if the events described are independent:

        \begin{itemize}
            \item[a.] When rolling a 6-sided die four times, let $E_{1}$
                be the event that the first two rolls sum to 7 and let $E_{2}$
                be the event that the last two rolls sum to 10. ~\\ \vspace{1cm}

            \item[b.] When choosing a committee of three dogs from a club of
                8 corgis and 12 labradors, let $E_{1}$ be the event that
                the committee has a labrador and let $E_{2}$ be the
                event that the committee has a corgi. ~\\ \vspace{1cm}
        \end{itemize}        
    \end{questionNOGRADE}

    \newpage

    \begin{intro}{The Product Rule, Theorem 3}
        If $E_{1}$ and $E_{2}$ are independent events in a given
        experiment, then the probability that both $E_{1}$ and $E_{2}$
        occur is the product of $Prob(E_{1}$ and $Prob(E_{2}$. That is,
        $$ Prob(E_{1} and E_{2}) = Prob(E_{1}) \cdot Prob(E_{2}) $$
        for independent events.
        \footnote{From Discrete Math by Ensley and Crawley, page 452}
    \end{intro}


    \stepcounter{question}
    \begin{questionNOGRADE}{\thequestion}
        Suppose I have a ``loaded" die for which the probability of a 6
        appearing is $\frac{1}{2}$, while the probability of each of the other faces
        appearing is $\frac{1}{10}$. What is the probability of getting a 5
        and then a 6 on two tosses of the loaded die?

        ~\\
        First identify $E_{1}$ and $E_{2}$. These events are independent,
        so you can use the Product Rule to find $Prob(E_{1} and E_{2})$.
    \end{questionNOGRADE}

    \newpage

        \begin{intro}{The General Product Rule, Theorem 4}
            If $E_{1}$ and $E_{2}$ are any events in a given experiment, then
            the probability that both $E_{1}$ and $E_{2}$ occur is given by

            $$Prob(E_{1} and E_{2}) = Prob(E_{2}) \cdot Prob(E_{1} | E_{2}) $$
            $$ = Prob(E_{1}) \cdot Prob(E_{2} | E_{1}) $$

            Note that if $E_{1}$ and $E_{2}$ are independent, then this says
            the same thing as Theorem 3.
            \footnote{From Discrete Math by Ensley and Crawley, page 453}
        \end{intro}

    \stepcounter{question}
    \begin{questionNOGRADE}{\thequestion}
        Two marbles are chosen from a bag containing three red, five white, and eight green marbles,
        so there are 16 total marbles.
        What is the probability that both are red?

        ~\\
        Here, the event $R_{1}$ is ``the first marble is red",
        and the event $R_{2}$ is ``the second marble is red".

        ~\\
        What is $Prob(R_{1})$? ~\\ \vspace{1cm}

        Since $R_{2}$ depends on $R_{1}$ occurring, after $R_{1}$
        occurs, there are 15 marbles left. One red marble
        has been selected, so there are 2 red marbles left.

        ~\\
        What is $Prob(R_{2} | R_{1})$? ~\\ \vspace{1cm}

        With this information, what is $Prob(R_{1} and R_{2})$?
    \end{questionNOGRADE}

    \newpage

    \stepcounter{question}
    \begin{questionNOGRADE}{\thequestion}
            Two marbles are chosen from a bag containing three red, five white, and eight green marbles,
            so there are 16 total marbles.
            What is the probability that one is white and one is green?

            ~\\
            Let's say we have the events
            $W_{1}$ (White first), $W_{2}$ (White second), $G_{1}$ (Green first), and
            $G_{2}$ (Green second), so we can get our result in two ways: with
            $(W_{1}, G_{2})$ \textbf{OR} with$(G_{1}, W_{2})$, so you can calculate
            the result as

            $$Prob(W_{1} and G_{2}) + Prob(G_{1} and W_{2})$$
    \end{questionNOGRADE}

}{
% KEY ------------------------------------ %

    \begin{enumerate}
        \item
            \begin{itemize}
                \item[a.]   When tossing a coin four times, no outcome can
                            consist of ``exactly two heads'' and also ``exactly three heads'';
                            hence, these two events are disjoint.
                            
                \item[b.]   When choosing four cards, if the four cards all have
                            the same value, then they cannot all have the same suit; hence, these
                            two events are disjoint.
                            
                \item[c.]   When choosing a committee of three people
                            from a club with 8 men and 12 women, there are many ways in which
                            the committee can include a woman and a man,
                            so these events are not disjoint.
            \end{itemize}

        \item
            \begin{itemize}
                \item[a.]   $$Prob(E_{1} or E_{2}) = Prob(E_{1}) + Prob(E_{2}) = \frac{1}{13} + \frac{1}{13} = \frac{2}{13}$$

                \item[b.]   $E_{1}$ is the outcomes where you get a diamond,
                            and $E_{2}$ is the outcomes where you get a black Jack, King, or Queen.
                            These sets are disjoint, so...
                            $$Prob(E_{1} or E_{2}) = Prob(E_{1}) + Prob(E_{2}) = \frac{1}{4} + \frac{6}{52} = \frac{19}{52}$$

                \item[c.]   $E_{1}$ is the outcomes where the card has an even numbered value,
                            and $E_{2}$ is the set of outcomes with a red Jack, King, or Queen.
                            These sets are disjoint, so...
                            $$Prob(E_{1} or E_{2}) = Prob(E_{1} + Prob(E_{2}) = \frac{5}{13} + \frac{6}{52} = \frac{1}{2}$$
            \end{itemize}

        \item
            \begin{itemize}
                \item[a.]   $n(S) = 36$
                \item[b.]   $E_1 = \{ (5, 1), (5, 2), (5, 3), (5, 4), (5, 5), (5, 6),
                            (6, 5), (4, 5), (3, 5), (2, 5), (1, 5) \}
                            \\ n(E_1) = 11$
                \item[c.]    $E_2 = \{ (2, 6), (6, 2), (3, 5), (5, 3), (4, 4) \} \\     n(E_2) = 5$

                \item[d.]   $n(E_1 \tab[0.5cm]AND\tab[0.5cm] E_2) = 2$

                \item[e.]   $\frac{11}{36} + \frac{5}{36} - \frac{2}{36} = \frac{7}{18}$
            \end{itemize}

        \item
            \begin{itemize}
                \item[a.]   The events are independent. The first two rolls
                            have no influence on the last two rolls.
                            
                \item[b.]   The events are not independent. The probability
                            of the committee having a corgi ($E_{2}$) is different
                            when $E_{1}$ occurs than it is when $E_{1}$ does not occur.
                            Specifically, if $E_{1}$ does not occur, then $E_{2}$ happens
                            for sure (i.e., its probability is 1), and if $E_{1}$ does occur,
                            then $E_{2}$ is not guaranteed to happen (i.e., probability is
                            less than 1).
            \end{itemize}

        \item   The events $E_{1}$, ``getting a 5 on the first toss",
                and $E_{2}$, ``getting a 6 on the second toss", are independent,
                so by the product rule,
                $Prob(E_{1} and E_{2}) = Prob(E_{1}) \cdot Prob(E_{2}) = \frac{1}{10} \cdot \frac{1}{2}
                = \frac{1}{20}$

        \item
            \begin{itemize}
                \item   $Prob(R_{1}) = \frac{3}{16}$
                \item   $Prob(R_{2}) = \frac{2}{15}$
                \item   $Prob(R_{1} and R_{2}) = Prob(R_{1}) \cdot Prob(R_{2} | R_{1})$ \\
                        $ = \frac{6}{240} = \frac{1}{40}$
            \end{itemize}

        \item   $W_{1}$ will be white as the first, and $G_{2}$ will be green as the second.

                $Prob(W_{1} and G_{1} + Prob(G_{1} and W_{2}) = \\
                    Prob(W_{1}) \cdot Prob(G_{2} | W_{1}) + Prob(G_{1}) \cdot Prob(W_{2}|G_{1}) \\
                    = \frac{5}{16} \cdot \frac{8}{15} + \frac{8}{16} \cdot \frac{5}{15} = \frac{1}{3}$
    \end{enumerate}

}

\input{BASE-FOOT}
