\input{BASE-HEAD}
\newcommand{\laClass}       {CS 211}
\newcommand{\laSemester}    {Spring 2018}
\newcommand{\laChapter}     {5.3}
\newcommand{\laType}        {Exercise}
\newcommand{\laPoints}      {5}
\newcommand{\laTitle}       {Combinations}
\newcommand{\laDate}        {Jan 23, 2018}
\setcounter{chapter}{5}
\setcounter{section}{1}
\addtocounter{section}{-1}
\newcounter{question}

\toggletrue{answerkey}
%\togglefalse{answerkey}

\input{BASE-HEADER}

\input{INSTRUCTIONS-EXERCISE}

\notonkey{
% ASSIGNMENT ------------------------------------ %

    \section{\laTitle}

    \subsection{Review}

    \begin{intro}{Permutation formula}
        A permutation is written as $P(n, r)$ where $n$ is the amount of
        items we have to choose from, and $r$ is the amount of items we
        are selecting. The formula for this is:

        $$ P(n, r) = \frac{n!}{(n-r)!} $$

        \paragraph{The Rule of Sums}
        
        \begin{quote}
            If we have A ways of doing something and B ways of doing
            another thing and we can not do both at the same time, then there are A + B ways to choose one of the actions.
            \footnote{From https://en.wikipedia.org/wiki/Rule\_of\_sum}
        \end{quote}

        \paragraph{The Rule of Complements}
        
        \begin{quote}
            If there are $x$ objects, and $y$ of those objects have a particular property,
            then the number of those objects that do \textbf{not} have that particular
            property is $x - y$.
            \footnote{From Discrete Mathematics, Ensley and Crawley, page 390}
        \end{quote}

        \paragraph{The Rule of Products}
        
        \begin{quote}
            If there are $a$ ways of doing something and $b$
            ways of doing another thing, then there are $a \cdot b$ ways of performing both actions.
            \footnote{From https://en.wikipedia.org/wiki/Rule\_of\_product}
        \end{quote}
        
    \end{intro}
    
    \subsection{Combinations and Permutations}

    \begin{intro}{Combinations}
        In mathematics, a combination is selection of items from a collection,
        such that (unlike permutations) the order of selection does not matter.
        (...) The number of $r$-combinations from a given set of $n$ elements
        is often denoted in elementary combinatorics texts by $C(n,r)$.
        \footnote{From https://en.wikipedia.org/wiki/Combinations}

        For a combination of length $r$ from a set of $n$ elements:
        $$C(n,r) = \frac{n!}{(n-r)! \cdot r!}$$

        (Note that the book uses $r$ and Wikipedia uses $k$.)
    \end{intro}
    

    \stepcounter{question}
    \begin{questionNOGRADE}{\thequestion}
        How many ways can you rearrange the letters in the word ``DOG''?
    \end{questionNOGRADE}
    
    \hrulefill
    
    \stepcounter{question}
    \begin{questionNOGRADE}{\thequestion}
        In how many ways can 10 children line up for lunch?
        \textit{Hint: Is this a combination problem or a permutation problem?
        Does order matter?}
    \end{questionNOGRADE}

    \hrulefill

    \stepcounter{question}
    \begin{questionNOGRADE}{\thequestion}
        In a class of 20 students, there are 7 IT majors and 13 CS majors.
        If 4 board positions had to be filled for the computer club,
        how many ways would there be to fill the positions with the given constraints?

        \begin{enumerate}
            \item[a.]   No constraints - any student can be on the board.
            \item[b.]   There must be exactly 2 IT students and 2 CS students on the board.
            \item[c.]   There must be \textit{at least} 2 IT students on the board.
        \end{enumerate}
    \end{questionNOGRADE}

    \newpage
    
    \stepcounter{question}
    \begin{questionNOGRADE}{\thequestion}
        Suppose we are going to receive a shipment of 50 games on floppy disk for our vintage game store.
        Each box of 50 generally has 3 defective floppies.
        For the shipment, we are going to select 5 games to feature in a display.

        \begin{enumerate}
            \item[a.]   How many total good floppies are there?
            \item[b.]   How many total bad floppies are there?
            \item[c.]   How many ways could we choose 5 games to feature?
            
            \item[d.]   How many ways contain \textit{no} defective floppies?
            
            \item[e.]   Using the Rule of Products, determine how many ways that contain all 3 defective floppies?
                    
            \item[f.]   Using the Rule of Sums and the Rule of Products, determine
                        how many ways contain \textit{at least one} defective floppy.

            \item[g.]   Using the Rule of Complements, determine
                        how many ways contain \textit{at least one} defective floppy.
                        The answer for (f) and (g) should match.
        \end{enumerate}
        
    \end{questionNOGRADE}

    \hrulefill

    \stepcounter{question}
    \begin{questionNOGRADE}{\thequestion}
        There's a bargain bin that has 5 PC games, 3 Playstation games, and 8 Xbox games.

        \begin{enumerate}
            \item[a.]   How many total games are there?
            
            \item[b.]   If you're grabbing 4 games to buy, you don't care about the order
                        that you pull the games out. How many ways can 4 games be selected?

            \item[c.]   How many ways can you select 4 games that are all for the same console?
                        \textit{(Hint: This means the Playstation games don't get counted.
                        Also, which Rule are you going to use to solve this?)}

            \item[d.]   How many selections of 4 games are there, where you have 2 for one platform,
                        and 2 for another platform?
                        \textit{(Hint: What Rules apply here?)}
        \end{enumerate}
    \end{questionNOGRADE}
    
}{
% KEY ------------------------------------ %
    \begin{itemize}
        \item[1.]   $P(3,3) = 6$
        
        \item[2.]   $P(10, 10) = 3628800$
        
        \item[3a.]  $C(20,4) = 4,845$
        \item[3b.]  $C(7,2) \cdot C(13,2) = 1,638$
        \item[3c.]  Two IT students, two CS students: $C(7,2) \cdot C(13,2) = 1,638$ \\
                    Three IT students, one CS student: $C(7,3) \cdot C(13,1) = 455$ \\
                    Four IT students, zero CS students: $C(7, 4) = 35$ \\
                    Result: $ 1638 + 455 + 35 = 2128 $

        \item[4a.]  There are 47 good floppies
        \item[4b.]  There are 3 bad floppies
        \item[4c.]  $C(50,5) = 2,118,760$
        \item[4d.]  $C(47,5) = 1,533,939$
        \item[4e.]  $C(3,3) \cdot C(47,2) = 1,081$
        \item[4f.]  $C(3,1) \cdot C(47,4) \tab + C(3,2) \cdot C(47,3) \tab + C(3,3) \cdot C(47,2) \tab = 584,821$ \\
        \item[4g.]  $C(50,5) - C(3,0) \cdot C(47,5) = 584,821$

        \item[5a.]  There are 16 total games.
        \item[5b.]  $C(16,4) = 1,820$
        \item[5c.]  $C(5,4) + C(8,4) = 75$
        \item[5d.]  $C(5,2) \cdot C(3,2) + C(5,2) \cdot C(8,2) + C(3,2) \cdot C(8,2) = 394$
    \end{itemize}
}

\input{BASE-FOOT}
